---
title: "Cluster Analysis-Whiskey Data"
output:
  pdf_document: default
  html_notebook: default
---
### Data Analytics and Visualizqation (Spring 2019)
### Instructor: Debopriya Ghosh

```{r}
# Reading the data
whiskies = read.csv("C:/Users/zhuwe/Desktop/Visualization/Dataset/whiskies.txt")
whiskies = whiskies[,-1]
sum(is.na(whiskies))  # no missing observations

```
```{r}
# generating a subset of the data that included only the 12 flavor variables, rescaled for comparability 
whiskies_k = scale(whiskies[,2:13])  # rescale selected vars for kmeans
head(whiskies_k)
```

```{r}
# applying k-means
ssPlot <- function(data, maxCluster = 9) {
  # Initialize within sum of squares
  SSw <- (nrow(data) - 1) * sum(apply(data, 2, var))
  SSw <- vector()
  for (i in 2:maxCluster) {
    SSw[i] <- sum(kmeans(data, centers = i)$withinss)
  }
  plot(1:maxCluster, SSw, type = "b", xlab = "Number of Clusters", ylab = "Within groups sum of squares")
}
ssPlot(whiskies_k)

```
Naturally, the within groups sum of squares decreases as we increase the number of clusters. However, there is a trend of diminishing marginal returns as we increase the number of clusters.Select the number of clusters based on the point at which the marginal return of adding one more cluster is less than was the marginal return for adding the clusters prior to that.
```{r}
fit <- kmeans(whiskies_k, 4)  # 4 cluster solution 
```

append cluster assignment
```{r}
whiskies <- data.frame(whiskies, fit$cluster)
whiskies$fit.cluster <- as.factor(whiskies$fit.cluster)
```
Cluster centers can inform on how taste profiles differ between clusters.
```{r}
fit$centers
```
Based on these centers, let us consider that David's choice for the full bodied, 
smoky and medicinal lies in cluster 4.
```{r}
subset(whiskies, fit.cluster == 4)
```
Identify the most representative whisky of each cluster by seeking out the observation closest to the center based on all 12 variables.
```{r}
whiskies_r <- whiskies[c(2:13, 17)]
# extract just flavor variables & cluster
candidates <- by(whiskies_r[-13], whiskies_r[13], function(data) {
  # we apply this function to observations for each level of fit.cluster
  dists <- sapply(data, function(x) (x - mean(x))^2)
  # for each variable, calc each observation's deviation from average of the
  # variable across observations
  dists <- rowSums(dists)
  # for each observation, sum the deviations across variables
  rownames(data)[dists == min(dists)]
  # obtain the row number of the smallest sum
})

candidates <- as.numeric(unlist(candidates))

whiskies[candidates, ]

```



