---
title: "DAV_hw03_More Remedies for Error Problems"
author: "Weijun Zhu"
date: "October 27, 2019"
output:
---
```{r}
library("printr")
library("faraway")
library("car")
library("MASS")
library("rgl")
library("MASS")
library("quantreg")
library("robustbase")
```


# Exercise 1
### a.Create factors, remove NAs
```{r}
data("uswages")
uswages$exper[uswages$exper < 0] = NA
uswages = na.omit(uswages)
```

### b.Process the data to make factors with named levels for race, smsa, and pt
```{r}
uswages$race = factor(uswages$race)
levels(uswages$race) = c("White", "Black")

uswages$smsa = factor(uswages$smsa)
levels(uswages$smsa) = c("No", "Yes")

uswages$pt = factor(uswages$pt)
levels(uswages$pt) = c("No", "Yes")
```

### c.Replace the indicator variables ne, mw, so, and we with one factor variable called region
```{r}
uswages = data.frame(uswages, region = 1 * uswages$ne + 2 * uswages$mw + 3 * 
    uswages$so + 4 * uswages$we)
uswages$region = factor(uswages$region)
levels(uswages$region) = c("ne", "mw", "so", "we")
uswages = subset(uswages, select = -c(ne:we))
```

### d.Remove all NAs
```{r}
uswages = na.omit(uswages)
```


# Exercise 2
### a.Show the summary of your final data
```{r}
summary(uswages)
```


# Exercise 3
### a.Compute the OLS fit to model log(wage)
```{r}
g = lm(wage ~ educ + exper + race + smsa + pt + region, uswages)
summary(g)
shapiro.test(residuals(g))
car::durbinWatsonTest(residuals(g))
car::vif(g)
par(mfrow=c(2,2))
plot(g)
```

### b.Perform Shapiro-Wilk test of Normality for the residuals, what is the conclusion?
```{r}
shapiro.test(residuals(g))
```
Conclusion: We reject the null hypothesis of normality for the residuals with level of significance 10% since the p-value is less than 0.10.

# Exercise 4
### a.Compute WLS fit to model log(wage) and weights = 1/(1+educ)
```{r}
g1 = lm(wage ~ educ + exper + race + smsa + pt + region, uswages)
g2 = lm(wage ~ educ + exper + race + smsa + pt + region, uswages,
         weight=1/(1+educ))
compareCoefs(g1, g2)
```

### b.Perform Shapiro-Wilk test of Normality for the residuals, what is the conclusion?
```{r}
shapiro.test(residuals(g2))
```
Conclusion: We reject the null hypothesis of normality for the residuals with level of significance 10% since the p-value is less than 0.10.


# Exercise 5
### a.Compute Robust fit to model log(wage), using Huber, Hampel, Bisquare, LTS and LAD
```{r}
# Huber
g2 = rlm(wage ~ educ + exper + race + smsa + pt + region, psi = psi.huber, uswages)

# Hampel
g4 = rlm(wage ~ educ + exper + race + smsa + pt + region, psi = psi.hampel, init="lts", maxit=100,
         uswages)
# Bisquare
g3 = rlm(wage ~ educ + exper + race + smsa + pt + region, psi = psi.bisquare, init="lts", maxit=100,
         uswages)
```

```{r}
# LTS
g5 = ltsReg(wage ~ educ + exper + race + smsa + pt + region, uswages)
g6 = ltsReg(wage ~ educ + exper + race + smsa + pt + region, uswages, nsamp="exact")
```

```{r}
# LAD
g7 = rq(wage ~ educ + exper + race + smsa + pt + region, data = uswages)
```


### b.Compare coefficients of the above fits using OLS, WLS, Huber, Hampel, Biquare, LTS, and LAD
```{r}

coefs = compareCoefs(g1, g2, g4, g3, se = FALSE)

colnames(coefs) = c("OLS", "Huber", "Hample", "Bisquare")

coefs
```

### c.Which would you recommend?
LTS is recommended since it has the best breakdown.






