---
title: "DAV_hw03_Model Selection and Model Validation"
author: "Weijun Zhu"
date: "October 27, 2019"
output: pdf_document
---
```{r}
library(car)
data(Salaries)
help(Salaries)
```

```{r}
Salaries$rownames = rownames(Salaries)
Salaries$rownames = NULL
Salaries.old = Salaries
Salaries = na.omit(Salaries)
levels(Salaries$discipline) = c("Theorectical", "Applied")
some(Salaries)
```

```{r}
Hitters = na.omit(Hitters)
```

```{r}
Salaries
```

# Exercise 1.
### Perform stepwise regression starting with the full model using all the predictors of salary.
```{r}
library(leaps)
regfit.full = regsubsets (salary ~ .,Salaries) 
summary(regfit.full)
```

# Exercise 2.
### Compare the coefficients of the stepwise model and the full model.
```{r}
coef(regfit.full ,6)
```

# Exercise 3.
### Which variable of variables did stepwise drop from the full model?
```{r}
coef(regfit.full ,6)
```
Conclusion: The best model of 6 variables show above.

# Exercise 4.
### Perform a cross-validation of the stepwise model.
```{r}
k = 3
set.seed(9) 
folds = sample(1:k,nrow(Salaries),replace=TRUE) 
cv.errors = matrix(NA,k,5, dimnames =list(NULL , paste(1:5)))
```

```{r}
predict.regsubsets = function (object ,newdata ,id,...){ 
  form=as.formula(object$call [[2]]) 
  mat=model.matrix(form,newdata) 
  coefi=coef(object ,id=id) 
  xvars=names(coefi) 
  mat[,xvars]%*%coefi 
  }
```

```{r}
for(j in 1:k){ 
  best.fit = regsubsets (salary ~ .,data = Salaries[folds!=j,], nvmax=5) 
  for(i in 1:5){ 
    pred = predict.regsubsets(best.fit ,Salaries[folds == j,],id = i) 
    cv.errors[ j,i] = mean((Salaries$salary[folds==j]-pred)^2) 
  } 
}
```

# Exercise 5.
### Compare the two models using the mse’s from the cross-validations with number of folds equal to 3. Which model gives the better mse?
```{r}
mean.cv.errors = apply(cv.errors,2,mean) 
mean.cv.errors
par(mfrow=c(1,1)) 
plot(mean.cv.errors,type = 'b')
```
Conclusion: We choose 3 variables model.


