---
title: 'Data Analysis and Visulation Homework #01'
output:
  pdf_document: default
  html_notebook: default
Name: Weijun Zhu;
---
```{r}
library(nycflights13)
library(tidyverse)
```

A. Sort flights to find the most delayed flights. Find the flights that left earliest
```{r}
# most delayed flights
flights %>%
  arrange(desc(dep_delay))

# Find the flights that left earliest
flights %>%
  arrange(dep_delay)

```


B. Which flights traveled the longest? Which traveled the shortest?
```{r}
# Travelled the longest
flights %>%
  arrange(desc(air_time))

# Travelled the shortest
flights %>%
  arrange(air_time)
```


C. Find flights that were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
flights %>%
  filter(dep_delay > 60, arr_delay - dep_delay > 30) %>%
  arrange(desc(dep_delay))

```

D. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?
```{r}
ave_delayed_flights = flights %>%
  group_by(year,month,day) %>%
  summarize(ave_dep_delay=mean(dep_delay,na.rm=TRUE),
            ave_arr_delay=mean(arr_delay,na.rm=TRUE))
#
cancelled_flights = flights %>%
  group_by(year,month,day) %>%
  summarize(num_cancelled_flights=sum(is.na(dep_delay)))
#
flights01 = merge(ave_delayed_flights,cancelled_flights)
#
ggplot(data=flights01) + 
  geom_point(mapping=aes(x=ave_arr_delay,y=num_cancelled_flights)) 
ggplot(data=flights01) + 
  geom_point(mapping=aes(x=ave_dep_delay,y=num_cancelled_flights))
```
Yes, the number of cancelled flights are related to the average arrival deplay and average departure deplay from the graphs. We can say the average deparure time and average arrival deplay increase, the number of cancelled flights increases.


E. What time of the day should you fly if you want to avoid delays as much as possible.
```{r}
flights %>%
  group_by(hour) %>%
  filter(!is.na(arr_delay)) %>%
  summarise(delay = mean(arr_delay)) %>%
  ggplot(aes(x = hour, y = delay)) + geom_point() 
# 
flights %>%
  group_by(hour) %>%
  filter(!is.na(dep_delay)) %>%
  summarise(delay = mean(dep_delay)) %>%
  ggplot(aes(x = hour, y = delay)) + geom_point()

```

```{r}
flights %>%
  group_by(hour) %>%
  filter(!is.na(arr_delay)) %>%
  summarise(delay = mean(arr_delay>0)) %>%
  ggplot(aes(x = hour, y = delay)) + geom_point() 
#
flights %>%
  group_by(hour) %>%
  filter(!is.na(dep_delay)) %>%
  summarise(delay = mean( dep_delay>0)) %>%
  ggplot(aes(x = hour, y = delay)) + geom_point()

```
We have four graphs of arrival delay and departure delay. From those four graphs, we can find if people take plane in the morning, between 5am - 10am, they will aviods delays as much as possible.


F. For each destination, compute the total minutes of delay. For each flight, compute the proportion of the total delay for its destination.
```{r}
# For each destination, compute total minutes of delay
flights %>%
  filter(arr_delay>0) %>%
  group_by(dest) %>%
  summarize(total_minutes=sum(arr_delay>0)) %>%
  arrange(desc(total_minutes))

```

```{r}
# For each flighs, compute the proportion of the total delay for its destination.
flights %>%
  filter(arr_delay>0) %>%
  group_by(dest) %>%
  mutate(arr_delay_prop = arr_delay / sum(arr_delay)) %>%
  select(carrier, flight, tailnum, origin, dest, arr_delay_prop) %>%
  arrange(desc(arr_delay_prop))

```
