---
title: "hw03-02"
author: "Weijun Zhu"
date: "November 12, 2019"
output: pdf_document
toc: yes
---
```{r}
library(tidyverse)
library(cluster)
library(fpc)
library(rgl)
```

```{r}
str(USArrests)
```

## 2a).
For each value of k from 2 to 10, run the k-means algorithm. For each run set the nstart
parameter to 20 so that the algorithm is run with 20 different random starting points. To make
sure all results are reproducible before starting run the command set.seed(3). This will make
all random numbers generated the same (this makes it easier to grade). Create an empty vector
ssw. For each k from 2 to 10, compute ssw=(total within ss)/(total ss). This is the ratio
of within-cluster sum-of-square distances to the total sum-of-square of all distances. By looking
at the plot decide what would be a good choice of k.
```{r}
n1 = 2
n = 10
km=kmTotss=kmBetweenss=kmWithinss=c()
for (i in n1:n){
  set.seed(3)
  km1<-kmeans(USArrests,i,iter.max=1000, nstart=20)
  kmTotss=c(kmTotss,km1$totss)
  kmWithinss=c(kmWithinss,sum(km1$withinss))
  kmBetweenss=c(kmBetweenss,km1$betweenss)
  km<-c(km,km1)
}
# plot(n1:n, kmBetweenss/sum(kmBetweenss), "b", col="blue", pch=20, xlab="k", ylab="SS err")
# lines(n1:n, kmWithinss/sum(kmWithinss), "b", col="orange", pch=20)
# lines(n1:n, kmTotss/sum(kmTotss), "b", col="red",pch=20)
plot(n1:n, kmWithinss/sum(kmTotss), "b", col="blue",pch=20,xlab="k", ylab="SS err")
```
Conclusion: 5 or 6 are good chance to be the number of clusters.

## 2b).
Set k, the number of clusters to 3. Run the k-means algorithm for k = 3 and with nstart=20.
Print the cluster of each state based on the result.
```{r}
model <- kmeans(USArrests,3,iter.max=1000, nstart=20)
clusplot(USArrests, model$cluster, color=TRUE, shade=TRUE, 
         labels=2, lines=0)
```
```{r}
library(factoextra)
fviz_cluster(model, data = USArrests)
```


## 2c).
Create a scatter plot with the X axis assault rate, and the Y axis as the murder rate. Color
those states in cluster 1 as blue, those in cluster 2 as orange, and those in cluster 3 as maroon.
Also, in addition to “dots" on the scatter plot, print the name of the state as well. For this check
documentation for the text function of R.
```{r}
USArrests['label'] <- model$cluster
```

```{r}
ggplot(USArrests, aes(x = USArrests$Assault, y = USArrests$Murder)) +
  geom_point(aes(color = factor(USArrests$label))) +
  scale_color_manual(breaks = c('1','2','3'), values = c('blue','orange','maroon'))
```

## 2d).
Repeat 2c) but using Rape on the X axis and Murder on the Y axis.
```{r}
ggplot(USArrests, aes(x = USArrests$Rape, y = USArrests$Murder)) +
  geom_point(aes(color = factor(USArrests$label))) +
  scale_color_manual(breaks = c('1','2','3'), values = c('blue','orange','maroon'))
```

## 2e).
Repeat 2c) but using Assault on the X axis and Rape on the Y axis.
```{r}
ggplot(USArrests, aes(x = USArrests$Assault, y = USArrests$Rape)) +
  geom_point(aes(color = factor(USArrests$label))) +
  scale_color_manual(breaks = c('1','2','3'), values = c('blue','orange','maroon'))
```

## 2f).
Repeat 2c) but using UrbanPop on the X axis and Murder on the Y axis.
```{r}
ggplot(USArrests, aes(x = USArrests$UrbanPop, y = USArrests$Rape)) +
  geom_point(aes(color = factor(USArrests$label))) +
  scale_color_manual(breaks = c('1','2','3'), values = c('blue','orange','maroon'))
```

## 2g).
Now load the rgl library. Plot a 3D graph with Murder on the X axis, Rape on the Y axis and
Assault on the Z axis. Use the same color code as the 2D plots for each cluster. Also add the
name of each state on each data point using text3d function (check the documentation).
```{r}
# library(rgl)
attach(USArrests)
plot3d(Murder,Rape,Assault,size=9,col=c('blue','orange','maroon')[USArrests$label])
text3d(USArrests$Murder,USArrests$Rape,USArrests$Assault,texts=row.names(USArrests))
detach(USArrests)
```

